---
title: "Lab 05: Harvesting research data"
author: "Logan Jacobs"
date: 2024-02-28
date-modified: today
format:
  html:
    toc: true
    toc_depth: 2
bibliography: [../bibliography.bib, ../packages.bib]
---

# Data description

The data I am downloading is the Switchboard Dialog Act Corpus. This Corpus is a set of conversations 

The data is in an archived file on the [LDC](https://catalog.ldc.upenn.edu/LDC97S62).

In this case, I am going to download, unarchive and save the contents of the corpus to the `data/original/swda` directory.

The format of the file, according to the README file is a list of annotated phone conversations, organized by 42 tags. The files themselves are .utt files, and are contained in a group of folders, which altogether constitutes the original unarchived .tar.gz file

The license of the file is denoted as "LDC User Agreement for Non-Members". Upon reading this file, the license generally describes that the database can be used for non-commercial linguistic education, research, and technology development, allowing me to use it for my purposes.

# Data collection

Below describes my data collection process for the data described above.

Before anything else, I will install the necessary packages. Here I am installing `dplyr`@dplyr for piping more efficiently and `fs`@fs for the file system.

```{r}
#| label: setup-packages
#| message: false

# Load packages
library(dplyr) #for piping
library(fs) #file system

```

The first step in the process is to try out using `download.file()` to acquire the archived file.

```{r}
#| label: download-archive-test

# set url to file address

url <- "https://catalog.ldc.upenn.edu/docs/LDC97S62/swb1_dialogact_annot.tar.gz"
data_dest <- "../data/original/swda.tar.gz"

download.file(url,data_dest)

```

Now the file is at `../data/original/swda.tar.gz`. Since it is a gzipped tarball file, I will have to unarchive it using `untar()` from base R.

```{r}
#| label: unarchive-test

untar(data_dest, exdir="../data/original/swda")
```

Upon completion of this process, the data will be in the appropriate directory: `../data/original/swda`.
